The Splashback Radius of Planck-SZ clusters

### 2019-03-13 Paper_v2/RSP_paper_new.tex
Fixed the COLLOSUS issue

The baryonic component of a galaxy cluster is dominated by the hot,
ionized  intra-cluster medium (ICM), which is gravitationally bound
within the cluster. The cosmic microwave background (CMB) photons
that pass through the cluster inverse Compton scatter off the hot
electrons and gain energy. This effect is known as the thermal
Sunyaev-Zeldovich (SZ) effect
\citep{sunyaev1970small,sunyaev1980velocity}. The effect has a
characteristic frequency dependence and results in an intensity
decrease below 220 GHz and an associated increase at higher
frequencies. The multiple frequency channels on the Planck satellite
allow a detection of galaxy clusters using the SZ effect
\citep{collaboration2016planck}. As part of the 2015 Data Release of
the Planck mission, the second Planck Catalogue of Sunyaev-Zeldovich
Sources (PSZ2) was made available to the community. The PSZ2 catalog
contains detections based on  three different techniques
\citep{ade2016planck}, and the union of these catalogs has in total
1653 galaxy clusters, of which 1203 clusters have been confirmed by
cross-matching to other galaxy clusters from external data sets.
\begin{table}
    \centering
    \caption{Comparison of the PSZ2 cluster catalog
    against the RedMaPPer cluster catalog used by
\citet{more2016detection}. The values of the mass estimates
$M_{\mathrm{500c}}$, $M_{\mathrm{200m}}$, redshifts $z$ and expected
splashback radii $r^{\mathrm{3D,theo}}_{\mathrm{sp}}$ represent the
catalog averages. The redshifts for both catalogs are given from the
survey. For the RedMaPPer catalog the $M_{\mathrm{200m}}$ mass
estimates were obtained from gravitational lensing and for the PSZ2
catalog the $M_{\mathrm{500c}}$ estimates were calculated from the
survey parameters using the scaling relation between the integrated
Compton Y-parameter $Y_{500c}$ and $M_{500c}$ as found by
\citet{ade2014planck}. The missing mass estimates as well as the
expected values of the splashback radii were calculated using the
Python package COLOSSUS \citep{diemer2017colossus}. The predictions
for the splashback radii $r^{\mathrm{3D,theo}}_{\mathrm{sp}}$ are
given in comoving units.}
    \label{tab:cluster_catalogs}
    \begin{tabular}{ccc}
    \hline
    & RedMaPPer & PSZ2 \\
    \hline
    $z$ & 0.24 & 0.177 \\
    \hline
    \# objects & 8643 & 596\\
    \hline
    $M_{\mathrm{500c}}$ [$h^{-1}10^{14} $M$_{\sun}$] & 0.9 & 3.0 \\
    \hline
    $M_{\mathrm{200m}}$ [$h^{-1}10^{14} $M$_{\sun}$] & 1.8 & 6.2\\
    \hline
    $r^{\mathrm{3D,theo}}_{\mathrm{sp}}$ [$h^{-1}$ Mpc] & 1.37 & 1.89 \\
    \hline
    \end{tabular}
\end{table}
\begin{table*}
    \centering
    \caption{Summary and comparison of the properties of the galaxy catalogs used in this work as well as the SDSS catalog used in \citet{more2016detection}.}
    \label{tab:galaxy_catalogs}
    \begin{tabular}{ccccc}
    \hline
    & SDSS & PS 21 & PS 21.5 & PS 22 \\
    \hline
    depth [mag] & 21.00 & 21.00 & 21.50 & 22.00\\
    \hline
    eff. area [deg$^2$] & $\approx$ 10,000 & 21,148 & 20,586 & 15,689\\
    \hline
    \# objects & 57,181,113 & 93,772,329 & 123,188,529 & 105,809,113 \\
    \hline
    objects/deg$^2$ & 5718 & 4434 & 5984 & 6744\\
    \hline
    \end{tabular}
\end{table*}

@@

We perform a preliminary comparison of the measurements with
expectations from cold dark matter models. The average halo mass of
the PSZ2 clusters as estimated from the Sunyaev-Zeldovich signal
is $M_{\rm 500c}=3.0\times10^{14}$ $\msunh$. We convert this mass
estimate to $M_{\rm 200m}=6.2\times10^{14}$ $\msunh$ using the average
concentration mass relation of halos \citep{Diemer2015} assuming an NFW profile
\citep{HuKravtsov:2003}. Given the average mass and redshift of our
cluster sample, we calculate the expected splashback radius to be
$1.89$ $\mpch$. We base this estimate on the fitting functions
presented in \citet{more2015splashback}. The splashback radius we
find for the three samples is consistent with this expectation,
although we can not rule out $\sim 15\%$ deviations in either directions,
given our large error bars.

@@

Our main findings can be summarized as follows:
\begin{itemize}
\item We detect a clear signature of a steepening feature in the
cross-correlation of Planck SZ clusters with Pan-STARRS
galaxies. The steepest logarithmic slopes that we find in our cross-correlation
signals are steeper than $-3$, and would hence be poorly fit by the
NFW profile. We associate this steepening with the splashback feature.
\item The location of the inferred splashback radius is $r_{\rm
sp}=1.85_{-0.30}^{+0.26}$ $\mpch$, which is consistent with expectations
from numerical simulations for
halos of an average mass $M_{\rm 500c}=3.0\times10^{14}\msunh$ at an
average redshift of $z=0.18$ in a collision-less dark matter Universe.
However, given the errors we cannot
currently rule out $\sim15\%$ deviations from these expectations.
\item We find that the location of the steepest slope does not
strongly depend on the magnitude of the galaxy samples we use, once
we go fainter than $M_i-5\log h=-19.44$.
\item By separately studying the cross-correlation of red and blue
galaxies with the clusters, we present evidence for the presence of
the splashback feature in both populations. The existence of the
splashback radius for the
star forming galaxy population could be of significance for the
models of satellite quenching in galaxy clusters. However, photometric
errors hinder a clean interpretation of the signal.
\end{itemize}

>>>>

The baryonic component of a galaxy cluster is dominated by the hot,
ionized  intra-cluster medium (ICM), which is gravitationally bound
within the cluster. The cosmic microwave background (CMB) photons
that pass through the cluster inverse Compton scatter off the hot
electrons and gain energy. This effect is known as the thermal
Sunyaev-Zeldovich (SZ) effect
\citep{sunyaev1970small,sunyaev1980velocity}. The effect has a
characteristic frequency dependence and results in an intensity
decrease below 220 GHz and an associated increase at higher
frequencies. The multiple frequency channels on the Planck satellite
allow a detection of galaxy clusters using the SZ effect
\citep{collaboration2016planck}. As part of the 2015 Data Release of
the Planck mission, the second Planck Catalogue of Sunyaev-Zeldovich
Sources (PSZ2) was made available to the community. The PSZ2 catalog
contains detections based on  three different techniques
\citep{ade2016planck}, and the union of these catalogs has in total
1653 galaxy clusters, of which 1203 clusters have been confirmed by
cross-matching to other galaxy clusters from external data sets.
\begin{table}
    \centering
    \caption{Comparison of the PSZ2 cluster catalog
    against the RedMaPPer cluster catalog used by
\citet{more2016detection}. The values of the mass estimates
$M_{\mathrm{500c}}$, $M_{\mathrm{200m}}$, redshifts $z$ and expected
splashback radii $r^{\mathrm{3D,theo}}_{\mathrm{sp}}$ represent the
catalog averages. The redshifts for both catalogs are given from the
survey. For the RedMaPPer catalog the $M_{\mathrm{200m}}$ mass
estimates were obtained from gravitational lensing and for the PSZ2
catalog the $M_{\mathrm{500c}}$ estimates were calculated from the
survey parameters using the scaling relation between the integrated
Compton Y-parameter $Y_{500c}$ and $M_{500c}$ as found by
\citet{ade2014planck}. The missing mass estimates as well as the
expected values of the splashback radii were calculated using the
Python package COLOSSUS \citep{diemer2017colossus}. The predictions
for the splashback radii $r^{\mathrm{3D,theo}}_{\mathrm{sp}}$ are
given in comoving units.}
    \label{tab:cluster_catalogs}
    \begin{tabular}{ccc}
    \hline
    & RedMaPPer & PSZ2 \\
    \hline
    $z$ & 0.24 & 0.177 \\
    \hline
    \# objects & 8643 & 596\\
    \hline
    $M_{\mathrm{500c}}$ [$h^{-1}10^{14} $M$_{\sun}$] & 1.0 & 3.0 \\
    \hline
    $M_{\mathrm{200m}}$ [$h^{-1}10^{14} $M$_{\sun}$] & 1.9 & 6.1\\
    \hline
    $r^{\mathrm{3D,theo}}_{\mathrm{sp}}$ [$h^{-1}$ Mpc] & 1.6 & 2.2 \\
    \hline
    \end{tabular}
\end{table}
\begin{table*}
    \centering
    \caption{Summary and comparison of the properties of the galaxy catalogs used in this work as well as the SDSS catalog used in \citet{more2016detection}.}
    \label{tab:galaxy_catalogs}
    \begin{tabular}{ccccc}
    \hline
    & SDSS & PS 21 & PS 21.5 & PS 22 \\
    \hline
    depth [mag] & 21.00 & 21.00 & 21.50 & 22.00\\
    \hline
    eff. area [deg$^2$] & $\approx$ 10,000 & 21,148 & 20,586 & 15,689\\
    \hline
    \# objects & 57,181,113 & 93,772,329 & 123,188,529 & 105,809,113 \\
    \hline
    objects/deg$^2$ & 5718 & 4434 & 5984 & 6744\\
    \hline
    \end{tabular}
\end{table*}

@@

We perform a preliminary comparison of the measurements with
expectations from cold dark matter models. The average halo mass of
the PSZ2 clusters as estimated from the Sunyaev-Zeldovich signal
is $M_{\rm 500c}=3.0\times10^{14}$ $\msunh$. We convert this mass
estimate to $M_{\rm 200m}=6.1\times10^{14}$ $\msunh$ using the average
concentration mass relation of halos \citep{Diemer2015} assuming an NFW profile
\citep{HuKravtsov:2003}. Given the average mass and redshift of our
cluster sample, we calculate the expected splashback radius to be
$2.2$ $\mpch$. We base this estimate on the fitting functions
presented in \citet{more2015splashback}. The splashback radius we
find for the three samples is marginally consistent with this expectation
but on the lower side. Given our large error bars further investigation is
warranted.

@@

Our main findings can be summarized as follows:
\begin{itemize}
\item We detect a clear signature of a steepening feature in the
cross-correlation of Planck SZ clusters with Pan-STARRS
galaxies. The steepest logarithmic slopes that we find in our cross-correlation
signals are steeper than $-3$, and would hence be poorly fit by the
NFW profile. We associate this steepening with the splashback feature.
\item The location of the inferred splashback radius is $r_{\rm
sp}=1.85_{-0.30}^{+0.26}$ $\mpch$, which is marginally consistent 
with expectations from numerical simulations for
halos of an average mass $M_{\rm 500c}=3.0\times10^{14}\msunh$ at an
average redshift of $z=0.18$ in a collision-less dark matter Universe.
However, given the errors further investigation is warranted.
\item We find that the location of the steepest slope does not
strongly depend on the magnitude of the galaxy samples we use, once
we go fainter than $M_i-5\log h=-19.44$.
\item By separately studying the cross-correlation of red and blue
galaxies with the clusters, we present evidence for the presence of
the splashback feature in both populations. The existence of the
splashback radius for the
star forming galaxy population could be of significance for the
models of satellite quenching in galaxy clusters. However, photometric
errors hinder a clean interpretation of the signal.
\end{itemize}
